---
title: "Practical Machine Learning"
author: "rtorlow"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
    toc: yes
    number_sections: yes
    css: docstyle.css
---

```{r prep, include = TRUE, eval=FALSE}

# load necessary R packages
library("knitr")
library("dplyr")
library("data.table")
library("caret")

# set seed
set.seed(5580)

# global options (can only be used after knitr was loaded)
opts_chunk$set(fig.align = "center", echo = FALSE,
               fig.retina = 10)

```

```{r setup, include = FALSE}
# load data
  load("08_assignment.rdata")

```

***
# Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset)


***
# Data input 

```{r,  eval=FALSE, echo=TRUE}
# raw data

Data <- as.data.frame(
  data.table::fread("pml-training.csv",
                    na.strings=c("NA", "", "#DIV/0!"),
                    strip.white=T, stringsAsFactors = TRUE))
```

```{r,  echo=TRUE}
dim(Data)
```

```{r,  eval=FALSE, echo=TRUE}
#  testing data for submission

TestingOrg <- as.data.frame(data.table::fread("pml-testing.csv", 
                  na.strings=c("NA", "", "#DIV/0!"), 
                  strip.white=T, 
                  stringsAsFactors = TRUE))
```

```{r,  echo=TRUE}
dim(TestingOrg)
```
***
# Data cleaning

```{r,eval=FALSE,  echo=TRUE}
#160 potential variables 
names(Data)

# remove irrelevant variables
# time stamps, username etc.
Data <- select(Data, -contains("timestamp"), -V1, -user_name, -contains("window"))

# lets find out the rate of missingness
isNA <- apply(Data, 2, function(x) { sum(is.na(x)) / nrow(Data) })
```

```{r, results='asis', echo=FALSE}

print(xtable::xtable(table(isNA)), 
      type = "html",include.rownames=TRUE)
```

```{r,eval=FALSE,  echo=TRUE}

# use only 53 variables which have no NA's (or any plausible rate of missingness)
Data <- Data[ , names(isNA[isNA==0])]
```
```{r,  echo=TRUE}
dim(Data)
```

```{r,eval=FALSE,  echo=TRUE}
# split into training and test data set
inTrain <- createDataPartition(Data$classe, p = 0.7, list = FALSE)
training <- Data[inTrain,]
testing <- Data[-inTrain,]
```

```{r,  echo=TRUE}
dim(training)
dim(testing)
```
***
# Train a random forest

Idea: Random forests are usually one of the two top performing algorithms (along with boosting). Random forests build lots of trees, and then average them to reduce the variance (Bagging: Bootstrap aggregating).

```{r,eval=FALSE,  echo=TRUE}
# get you a coffee... it takes time
model <- train(classe ~ ., data = training, model="rf", trControl=ctrl)
pred <- predict(model, newdata = testing)
```


## Out-of-sample-error
```{r,  echo=TRUE}
sum(pred == testing$classe) / length(pred) 
# 98% of the test data set is predicted right

# the out-of-sample-error is:
1 - sum(pred == testing$classe) / length(pred) 
```

```{r,eval=FALSE,  echo=TRUE}
# we can table which classe is predicted right / wrong
confusionMatrix(testing$classe, pred)$table
```
```{r,  echo=F}
print(a)
```

***
## Which variables are most important in this model?
```{r,eval=FALSE,  echo=TRUE}
top10 <- varImp(model)$importance
top10 <- mutate(top10, vars = row.names(top10)) %>%
  arrange(desc(Overall))

top10 <- top10$vars[1:10]
```
```{r,  echo=TRUE}
top10
```

***
# Train a smaller random forest

```{r,eval=FALSE,  echo=TRUE}

smallData   <- training[, c( "classe", top10)]
smallModel  <- train(classe ~ ., data = smallData, model = "rf", trControl=ctrl)
smallPred   <- predict(smallModel, newdata = testing)
```

## Out-of-sample-error (small tree)
```{r,  echo=TRUE}

sum(smallPred == testing$classe) / length(smallPred)

# Out-of-sample-error for model with 10 variables
1- sum(smallPred == testing$classe) / length(smallPred)
```

```{r,eval=FALSE,  echo=TRUE}
# Finally, let's check which classe is predicted right / wrong
confusionMatrix(testing$classe, smallPred)$table
```
```{r,  echo=F}
b
```

***
# Prediction for submission
```{r,eval=FALSE,  echo=TRUE}
answers = rep("A", 20)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

predict(model, newdata=TestingOrg) # use the orignal model
answers <- predict(model, TestingOrg)

pml_write_files(answers)
```

```{r,  echo=TRUE}
answers
```

